### 2021-09-15 papers.Rmd
first attempt

---
title: "papers"
site: distill::distill_article

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

aside_center <- function(text) {
  return(aside(center(list(text))))
}
```

## 2021

Lim, M. C., **Parsons, S.**, Goglio, A., Fox, E. (2021). Anxiety, stress, and binge eating behaviours in adolescence: a prospective approach. *Journal of Eating Disorders, 9*. [doi.org/10.1186/s40337-021-00444-2](https://doi.org/10.1186/s40337-021-00444-2)

**Parsons, S.**, (2021). splithalf: robust estimates of split half reliability. *Journal of Open Source Software, 6*(60), 3041, [doi.org/10.21105/joss.03041](https://doi.org/10.21105/joss.03041)

<div data-badge-type="donut" data-doi="10.21105/joss.03041" data-hide-no-mentions="true" class="altmetric-embed"></div>

<script type='text/javascript' src='https://d1bxh8uas1mnw7.cloudfront.net/assets/embed.js'></script>

**Parsons, S.**, Booth, C., Songco, A., & Fox, E. (2021). Emotional information-processing correlates of mental health in adolescence: A network analysis approach. *Cognition and Emotion* \\ [doi.org/10.1080/02699931.2021.1915752](https://doi.org/10.1080/02699931.2021.1915752). Preprint: [psyarxiv.com/8ygn](https://psyarxiv.com/8ygnh), data and code: [osf.io/mn5ek/](https://osf.io/mn5ek/)

Demkowicz, O., Panayiotou, M., **Parsons, S.**, Feltham, A., Arseneault, L., Ingram, B., Patalay, P., Edge D.,, Pierce, M., Creswell, C., Victor, C., O'Connor, R., & Qualter, P. (2021). Looking back to move forward: Reflections on the strengths and challenges of the COVID-19 UK mental health research response. 
*Frontiers Psychiatry*. [10.3389/fpsyt.2021.622562](https://doi.org/10.3389/fpsyt.2021.622562)

## 2020

Songco, A., Booth, C., Spiegler, O., **Parsons, S.**, & Fox, E. (2020). Anxiety and depressive symptom trajectories in adolescence and the co-occurring development of cognitive biases: Evidence from the CogBIAS longitudinal study. *Journal of Abnormal Child Psychology, 48*, 1617–1633, DOI: [10.1007/s10802-020-00694-9](https://doi.org/10.1007/s10802-020-00694-9)

Booth, C., Songco, A., **Parsons, S.**, & Fox, E. (2020). Cognitive mechanisms predicting resilient functioning in adolescence: Evidence from the CogBIAS longitudinal study. *Development and Psychopathology*, 1-9. DOI: [10.1017/S0954579420000668](https://doi.org/10.1017/S0954579420000668)

Dodd, H., Rayson, H., Ryan, Z., Bishop, C., **Parsons, S.**, Stuijfzand, B. (2020). Trajectories of anxiety when children start school: The role of behavioral inhibition and attention bias to angry and happy faces
*Journal of Abnormal Psychology*. DOI: [10.1037/abn0000623](https://doi.org/10.1037/abn0000623)

**Parsons, S.** (2020). Does attention bias modification induce structural brain changes? A commentary on Abend et al. (2019). *Biological Psychology, 152*. DOI: [10.1016/j.biopsycho.2020.107866](https://www.sciencedirect.com/science/article/pii/S0301051120300260). Preprint: [psyarxiv.com/4ze6h](https://psyarxiv.com/4ze6h/)

## 2019

Herbert, M., Notebaert, L., **Parsons, S.**, Fox, E., & MacLeod, C. (2019). The Effect of Varying Danger Controllability on Attention to Threat Messages. *Applied Cognitive Psychology, 34*(2). DOI: [10.1002/acp.3628](https://onlinelibrary.wiley.com/doi/abs/10.1002/acp.3628)[10.1002/acp.3628]

**Parsons, S.**, Kruijt, A-W., & Fox, E. (2019). Psychological Science needs a standard practice of reporting the reliability of cognitive behavioural measurements. *Advances in Methods and Practices in Psychological Science, 2,*(4), 378-395. DOI: [doi.org/10.1177/2515245919879695](https://doi.org/10.1177/2515245919879695). Preprint: [psyarxiv.com/6ka9z](https://psyarxiv.com/6ka9z)

Booth, C., Songco, A., **Parsons, S.**, Heathcote, L.C., & Fox, E. (2019). The CogBIAS longitudinal study of adolescence: Cohort profile and stability and change in measures across three waves. *BMC Psychology, 7*, 73. DOI: [10.1186/s40359-019-0342-8](https://bmcpsychology.biomedcentral.com/articles/10.1186/s40359-019-0342-8)

Crüwell, S., van Doorn, J., Etz, A., Makel, M.C., Moshontz, H., Niebaum, J., Orben, A., **Parsons, S.**, & Schulte-Mecklenbeck, M. (2019). 7 Easy Steps to Open Science: An annotated reading list. *Zeitschrift für Psychologie, 227*, 237-248. DOI: [doi.org/10.1027/2151-2604/a000387](https://doi.org/10.1027/2151-2604/a000387). Preprint: [psyarxiv.com/cfzyx](https://psyarxiv.com/cfzyx)

Kruijt, A-W., **Parsons, S.**, & Fox, E. (2019). A meta-analysis of bias at baseline in RCTs of attention bias modification: no evidence for bias towards threat in clinical anxiety and PTSD. *Journal of Abnormal Psychology, 128*(6), 563-573. DOI: 10.1037/abn0000406. Preprint DOI: [10.31234/osf.io/rfjup](10.31234/osf.io/rfjup)

## 2018

**Parsons, S.** (2018). Moving forward with questions of process and procedure in cognitive bias modification research: three points of consideration. *Published as an eLetter on* [10.1192/bjp.2018.43](https://doi.org/10.1192/bjp.2018.43). Preprint: [psyarxiv.com/k3vxc](https://psyarxiv.com/k3vxc)

Notebaert, L., Georgiades, J., Herbert, M., Grafton, B., **Parsons, S.**, Fox, E., & MacLeod, C. (2018). Trait anxiety and the alignment of attentional bias with controllability of danger. *Psychological Research*. DOI: [doi.org/10.1007/s00426-018-1081-9](https://doi.org/10.1007/s00426-018-1081-9)

## 2017

Lakens, et al. (2017). Justify Your Alpha: A Response to “Redefine Statistical Significance”. *Nature Human Behaviour, 2*, 168-171. DOI: [10.1038/s41562-018-0311-x](https://www.nature.com/articles/s41562-018-0311-x). Preprint: [psyarxiv.com/9s3y6/](https://psyarxiv.com/9s3y6/)

Booth, C., Songco, A., **Parsons, S.**, Heathcote, L., Vincent, J., Keers, R., & Fox, E. (2017). The CogBIAS longitudinal study protocol: cognitive and genetic factors influencing psychological functioning in adolescence. *BMC Psychology, 5*(1). DOI: [doi.org/10.1186/s40359-017-0210-3](https://doi.org/10.1186/s40359-017-0210-3)

## 2016

**Parsons, S**, Kruijt, A-W., & Fox, E. (2016). A cognitive model of psychological resilience. *Journal of Experimental Psychopathology, 7*(3), 296-310. DOI: http://dx.doi.org/10.5127/jep.053415