### 2021-09-29 index.Rmd

---
title: "Sam Parsons"
image: "SamProfile.jpg"
preview: "SamProfile.jpg"
twitter:
  site: "@sam_d_parsons"
links:
  - label: '<i class="fab fa-twitter"></i> Twitter'
    url: "https://twitter.com/sam_d_parsons"
  - label: '<i class="fab fa-github"></i> GitHub'
    url: "https://github.com/sdparsons"
  - label: '<i class="fas fa-envelope"></i> Email'
    url: "mailto:sam.parsons@psy.ox.ac.uk"
  - label: '<i class="ai ai-orcid"></i> ORCID'
    url: "https://orcid.org/0000-0002-7048-4093"
  - label: '<i class="ai ai-publons"></i> Publons'
    url: "https://publons.com/researcher/1351322/sam-parsons/"
  - label: '<i class="fab fa-youtube"></i> YouTube'
    url: "https://www.youtube.com/channel/UC3C_BRgKQNqZH2P2R5U2cfw"
  - label: '<i class="ai ai-osf"></i> OSF'
    url: "https://osf.io/k7akr/"
  - label: '<i class="ai ai-cv"></i> CV'
    url: "Sam_Parsons_cv.pdf"
output: 
  postcards::jolla
site: distill::distill_website
---

# [<i class="fab fa-twitter"></i>](https://twitter.com/sam_d_parsons) [<i class="ai ai-google-scholar"></i>](https://scholar.google.co.uk/citations?user=kE37Xy4AAAAJ&hl=en) [<i class="ai ai-orcid"></i>](https://orcid.org/0000-0002-7048-4093) [<i class="ai ai-osf"></i>](https://osf.io/k7akr/) [<i class="ai ai-researchgate"></i>](https://www.researchgate.net/profile/Sam_Parsons4) [<i class="fab fa-github"></i>](https://github.com/sdparsons) [<i class="ai ai-publons"></i>](https://publons.com/researcher/1351322/sam-parsons/) [<i class="fab fa-youtube"></i>](https://www.youtube.com/channel/UC3C_BRgKQNqZH2P2R5U2cfw) [<i class="fas fa-envelope-square"></i>](mailto:sam.parsons@psy.ox.ac.uk) [<i class="ai ai-cv"></i>](Sam_Parsons_cv.pdf)

I’m a [Radboud Excellence Initiative](https://www.ru.nl/excellence/) Postdoctoral Research Fellow. I work in the [Lifespan Cognitive Dynamics Lab](https://www.rogierkievit.com/) based at the [Donders Institute](https://www.ru.nl/donders/). I did my DPhil in Experimental Psychology at the University of Oxford (a DPhil is a PhD but of course they need to call it something else)

Just in case you’re here for the cat, this ginger monster is called Beau.

Radboud project

Open Scholarship - ReproducibiliTea and FORRT

I try to do my own small bit for the improvement of psychological science. Sometimes this is promoting open and reproducible research practices; others it is frustrating others by highlighting the importance of routinely estimating and reporting the reliability of our measurements. Mostly, I’d just like us all to get along and do good research.

Previous projects & talks

Seriousness - not a huge fan

I am happy to be contacted to discuss reliability of cognitive measures - particularly if you have questions about my splithalf package and need assistance with it.

(I try to keep this website up to date, however it may fall slightly behind. I do keep my [CV](Sam_Parsons_cv.pdf) up to date, so please do refer to that)

---------------------------

## splithalf

Estimate the internal consistency of your tasks with a permutation based split-half reliability approach.

Latest version: 0.8.1 “Rum Ham”

[<i class="fab fa-github"></i> Github documentation](https://github.com/sdparsons/splithalf)

Now published in the [Journal of Open Source Software](https://joss.theoj.org/papers/10.21105/joss.03041)

![](splithalf.png)